Evolution of \\ ice sheet geometry \\ using Stokes dynamics

### 2021-06-02 talk/talk.tex
A bit more.

\vspace{-1mm}
\begin{itemize}
\item a non-shallow glacier interacting with the climate
    \begin{itemize}
    \item only the simplest such problem (nonsliding, isothermal)
    \item re~\emph{steady}: solve this and you can handle implicit steps too
    \end{itemize}
\item data given on fixed $\Omega \subset \RR^2$:
    \begin{itemize}
    \item climatic mass balance $a(x,y)$ \hfill $\gets$ \emph{steady climate}
    \item bed elevation $b(x,y)$
    \end{itemize}
\item determine: surface elevation $s(x,y)$ on $\Omega$ and domain $\Lambda_s \subset \RR^3$
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item $\bn_s = \left<-s_x,-s_y,1\right>$  \quad (normal to ice surface; upward)
\item \emph{on} the ice surface, in steady-state there is balance (SKE) between motion and climatic mass balance
\item \emph{off} the ice we know $a\le 0$ in steady state
\item for all of this talk: $s=b$ where ice is not present
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item \alert{a nonlinear complementarity problem (NCP) for $s$ coupled to Stokes problem for $\bu,p$:}

@@

\begin{itemize}
\item using $\Phi$ gives clean NCP over $\Omega$:
\begin{align*}
s - b &\ge 0 \\
- \Phi(s) - a &\ge 0 \\
(s - b) (- \Phi(s) - a) &= 0
\end{align*}

\item well-known fact: \quad  NCP (strong form) $\leftrightarrow$ VI (weak form)
\item find $s \in \mathcal{K} = \{r \ge b\}$ so that
    $$\hspace{10mm} F(s)[r - s] \ge \ip{a}{r-s} \hspace{10mm} (\ast)$$
for all $r \in \mathcal{K}$, where
    $$F(s)[w] = - \int_\Omega \Phi(s)\, w \,dx dy$$
    \begin{itemize}
    \item whether VI problem $(\ast)$ is well-posed is open, but the SIA version has an existence proof (Jouvet \& Bueler \cite{JouvetBueler2012})
    \end{itemize}
\end{itemize}

@@

\begin{block}{goal}
given triangulation $\mathcal{T}$ of $\Omega$, and $P_1$ FE space $\mathcal{V}^h \subset W^{1,\qq}(\Omega)$, solve SIGP VI

$$F(s^h)[r^h - s^h] \ge \ip{a}{r^h-s^h}$$

for admissible $s^h$ using solutions of Glen-Stokes problem $F_{\Lambda_{s^h}}(\bu^h,p^h)=0$ to evaluate $\Phi(s)=\bu|_s\cdot\bn_s$ and $F(s)[w] =- \int_\Omega \Phi(s)\, w $
\end{block}

\begin{itemize}
\item seeking a solver which is:
    \begin{itemize}
    \item \alert{robust}
    \item \alert{optimal} over $m$ degrees of freedom in $\mathcal{V}^h$: work is $O(m)$
    \end{itemize}
\item optimal requires multigrid
\end{itemize}
\end{frame}

>>>>

\vspace{-1mm}
\begin{itemize}
\item a glacier which is in balance with its climate
    \begin{itemize}
    \item the simplest such problem (grounded, nonsliding, isothermal)
    \item re~\emph{steady}: solve this and you can handle implicit steps too
    \end{itemize}
\item steady climate data given on fixed $\Omega \subset \RR^2$:
    \begin{itemize}
    \item climatic mass balance $a(x,y)$ \hfill $\gets$ \emph{$a>0$ for snowing}
    \item bed elevation $b(x,y)$
    \end{itemize}
\item determine: surface elevation $s(x,y)$ on $\Omega$, domain $\Lambda_s \subset \RR^3$
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item $\bn_s = \left<-s_x,-s_y,1\right>$  \quad (normal to ice surface; upward)
\item \emph{on} the ice surface, in steady-state there is balance (SKE) between motion and climatic mass balance
\item \emph{off} the ice we know $a\le 0$ in steady state
\item for all of this talk:
    \begin{itemize}
    \item $s$ is assumed to be well-defined (no overhangs)
    \item $s=b$ where ice is not present, so $s$ is defined on all of $\Omega$
    \end{itemize}
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item the SIGP is a \alert{nonlinear complementarity problem (NCP) for $s$ which is coupled to a Stokes problem for $\bu,p$:}

@@

\begin{itemize}
\item using $\Phi$ gives a cleaner NCP over $\Omega$:
\begin{align*}
s - b &\ge 0 \\
- \Phi(s) - a &\ge 0 \\
(s - b) (- \Phi(s) - a) &= 0
\end{align*}

\item well-known fact: \quad  NCP (strong form) $\leftrightarrow$ VI (weak form)
\item VI weak form: find $s \in \mathcal{K} = \{r \ge b\}$ so that
    $$\hspace{10mm} F(s)[r - s] \ge \ip{a}{r-s}$$
for all $r \in \mathcal{K}$, where $F(s)[w] = - \int_\Omega \Phi(s)\, w \,dx dy$

\medskip
    \begin{itemize}
    \item whether this VI is well-posed is open, but the SIA version has an existence proof (Jouvet \& Bueler \cite{JouvetBueler2012})
    \end{itemize}
\end{itemize}

@@

\begin{block}{numerical goal}
Given triangulation $\mathcal{T}$ of $\Omega$ and $P_1$ space $\mathcal{V}^h \subset W^{1,\qq}(\Omega)$.  Solve VI

$$F(s^h)[r^h - s^h] \ge \ip{a}{r^h-s^h} \qquad \text{for all $r^h \in \mathcal{K}^h$}$$

to give admissible $s^h \in \mathcal{K}^h$.  Use the solution of the Stokes problem $F_{\Lambda_{s^h}}(\bu^h,p^h)=0$ to evaluate $\Phi(s^h)=\bu|_{s^h}\cdot\bn_{s^h}$ (and thus $F(s^h)$).
\end{block}

\begin{itemize}
\item seeking a solver which is:
    \begin{itemize}
    \item \alert{robust}
    \item \alert{optimal} over $m$ degrees of freedom in $\mathcal{V}^h$: work is $O(m)$
    \end{itemize}
\item optimal requires multigrid
\item<2> \emph{for now these are aspirations only, but let's give it a shot}
\end{itemize}
\end{frame}
